
# Analyse basique de la structure des publis


```{r include = FALSE}
# setup

library(dplyr)
library(ggplot2)

setwd(paste0(Sys.getenv('CS_HOME'),'/Misc/AnalyseGeocites/Models'))


# authors
authors <- as.tbl(read.csv('data/authors.csv',sep=";",stringsAsFactors = FALSE))

# publis
publis <- as.tbl(read.csv('data/publications.csv',sep=";",stringsAsFactors = FALSE))

# filter on type publis
publis = publis[publis$typepubli%in%c("article_revue","proceedings"),]

# attributes
properties <- as.tbl(read.csv('data/properties.csv',sep=";",stringsAsFactors = FALSE))

axes = properties$nameshort[properties$type=='axe']
teams = properties$nameshort[properties$type=='equipe']
allyears = setdiff(unique(publis$year),c("2017"))

```


## Stats basique

### Publis hors axe
```{r} 
100*length(which(publis$axe=='hors_axe'))/nrow(publis)
```

### Distrib of interaxe collabs

```{r}
axe_num = sapply(publis$axe,function(s){length(strsplit(s,',')[[1]])})
hist(axe_num,breaks=5)
```


### Interequipes

```{r}
eq_num = sapply(publis$equipe,function(s){length(strsplit(s,',')[[1]])})
hist(eq_num,breaks=5)
```


Interéquipe dans le temps ?





## Interne/externe

Publis with one author
```{r}
authornum = sapply(publis$auteurs,function(s){length(strsplit(s,',')[[1]])})
length(which(authornum==1))/length(authornum)
hist(authornum,breaks=50,xlab="number of authors",main="")
```

Internal rate for each publi with more than one author :
```{r}
isumr = authors$umr;names(isumr)<-authors$id

internrate<-function(teams){
  res=c()
  for(team in teams){
    res=append(res,sapply(publis$auteurs[publis$equipe==team],function(s){
      auths = strsplit(s,',')[[1]];
      currentumr = isumr[auths];
      if(length(currentumr)>1){return(sum(currentumr)/length(currentumr))}else{return(NA)}
    })
    )
  }
  return(res)
}
hist(internrate(c("CRIA","EHGO","PARIS")),breaks=50)
hist(internrate(c("CRIA")),breaks=50)
hist(internrate(c("EHGO")),breaks=50)
hist(internrate(c("PARIS")),breaks=50)
#mean(internrate,na.rm = T)
```




## Export graph axes gexf

Graph construction function
```{r include = FALSE}
library(rgexf)
library(igraph)

exportGraph <- function(nodesid,linkcolumn,out,type=c("gexf","igraph-year","igraph")){
  # id,label
  nodes = data.frame(id=nodesid,label=nodesid)
  edges=matrix(rep("",2*length(nodesid)^2),nrow =length(nodesid)^2);k=1
  for(a1 in 1:length(nodesid)){show(a1)
    for(a2 in 1:length(nodesid)){edges[k,1]=nodesid[a1];edges[k,2]=nodesid[a2];k=k+1}}
  # edges : id1,id2
  # weights -> by publi number ; or by author ? -> the first is the second but weighted
  counts=list();
  #for(a1 in nodesid){show(a1)
  #  for(a2 in nodesid){for(y in unique(publis$year)){counts[[paste0(a1,a2,y)]]=0}}}
  for(i in 1:nrow(publis)){
    show(i)
    raxes = as.character(publis[i,linkcolumn]);
    currentaxes = strsplit(raxes,',')[[1]]
    if(length(currentaxes)>1){
      for(a1 in currentaxes){for(a2 in currentaxes){
        key=paste0(a1,a2,publis$year[i])
        if(!(key%in%names(counts))){counts[[key]]=0}
        counts[[key]]=counts[[key]]+1
        }}
    }
  }
  
  weights=c();years=c();j=1
  alledges=data.frame()
  for(y in unique(publis$year)){
    show(y)
    j=1;inds=c();currentweights=c()
    for(a1 in nodesid){
      show(a1)
      for(a2 in nodesid){
      #show(counts[[paste0(a1,a2,y)]])
        key=paste0(a1,a2,y)
      if(key%in%names(counts)){
        inds=append(inds,j);currentweights = append(currentweights,counts[[paste0(a1,a2,y)]])
        years=append(years,y)
      }
      j=j+1
    }}
    alledges=rbind(alledges,edges[inds,])
    weights=append(weights,currentweights)
    
    if(type=="igraph-year"){
      g = graph_from_data_frame(cbind(edges[inds,],currentweights),directed = FALSE)
      save(g,file = paste0('res/graph_',y,'.RData'))
    }
    
  }
  
  if(type=="gexf"){
    write.gexf(nodes=nodes,
               edges=alledges,
               edgesWeight=weights,
               edgeDynamic=data.frame(years,years+0.99),
               output=out
    )
  }
}

```



```{r}
exportGraph(axes,8,'res/axes.gexf',type="gexf")

```



## co-autorship in time

```{r}
exportGraph(authors$id,3,'res/coauthor.gexf',type="gexf")
```




## Evolution of modularity in time ?

-> for co-authorship network

Define modularity functions
```{r include = FALSE}
library(Matrix)

overlappingmodularity <- function(probas,adjacency){#,linkfun=function(p1,p2){return(p1*p2)}){
  show(paste0('Computing overlapping modularity : dim(probas)=',dim(probas)[1],' ',dim(probas)[2],' ; dim(adjacency)=',dim(adjacency)[1],' ',dim(adjacency)[2]))
  m = sum(adjacency)
  n=nrow(probas)
  kout=rowSums(adjacency)
  kin=colSums(adjacency)
  res=0
  for(c in 1:ncol(probas)){
    if(sum(probas[,c])>0){
      if(c%%100==0){show(c/ncol(probas))}
      a1 = Diagonal(x=probas[,c])%*%adjacency%*%Diagonal(x=probas[,c])
      a2 = sum(kout*probas[,c])*sum(kin*probas[,c])*((sum(probas[,c])/n)^2)/m
      res = res + sum(a1) - a2
      rm(a1);gc() # loose time to call gc at each step ?
    }
  }
  return(res/m)
}

directedmodularity<-function(membership,adjacency){
  # sum([A_ij - k_iout k_j in/m ]\delta (c_i,c_j))
  # 
  #deltac = sparseMatrix(1:nrow(adjacency),1:ncol(adjacency),x=0)
  #inds=c()
  #for(c in unique(membership)){inds = append(inds,which(membership==c))}
  m=sum(adjacency)
  kout=rowSums(adjacency);kin=colSums(adjacency)
  res = 0;k=length(unique(membership))
  for(c in unique(membership)){
    #if(c%%100==0){show(c/k)}
    inds=which(membership==c)
    res = res + sum(adjacency[inds,inds]) - sum(kin[inds])*sum(kout[inds])/m 
    gc()
  }
  return(res/m)
}

```

Compute yearly co-authorship networks
```{r include = FALSE}
exportGraph(authors$id,3,'',type='igraph-year')
```

get classes proba matrix
```{r include = FALSE}
authorsProbas<-function(authorsid,colid,classesnames,years){
  res = matrix(0,length(authorsid),length(classesnames))
  rownames(res)<-authorsid;colnames(res)<-classesnames
  currentpublis=publis[publis$year%in%years,]
  for(i in 1:nrow(currentpublis)){
    raxes = as.character(currentpublis[i,colid]);
    currentaxes = strsplit(raxes,',')[[1]]
    rauthors = as.character(currentpublis[i,3]);
    currentauthors = strsplit(rauthors,',')[[1]]
    for(currentauthor in currentauthors){
    for(currentaxe in currentaxes){
      if(currentauthor%in%authorsid&currentaxe%in%classesnames){
        res[currentauthor,currentaxe]=res[currentauthor,currentaxe]+1
      }
    }} 
  }
  return(t(apply(res,1,function(r){r/sum(r)})))
}
```


Modularity in time
```{r include = TRUE}

mods= c();years=c();types=c()
for(year in allyears){
  show(year)
  # load graph
  load(paste0('res/graph_',year,'.RData'))
  adjacency = as.matrix(as_adjacency_matrix(g,type = "both"))
  axesprobas = authorsProbas(V(g)$name,8,axes,c(year))
  teamprobas = authorsProbas(V(g)$name,6,teams,c(year))
  mods=append(mods,overlappingmodularity(axesprobas,adjacency))
  mods=append(mods,overlappingmodularity(teamprobas,adjacency))
  years=append(years,c(year,year));types=append(types,c("axes","equipes"))
}
```

plot
```{r include = TRUE}
g=ggplot(data.frame(mod=mods,year=years,type=types),aes(x=year,y=mod,colour=type,group=type))
g+geom_point()+geom_line()
```


## Clustering coefficients (easier to interpret than modularity)

```{r include = TRUE}
ccoefs= c();years=c();types=c();ccoefnames=c()
for(year in allyears){
  show(year)
  # load graph
  load(paste0('res/graph_',year,'.RData'))
  adjacency = as.matrix(as_adjacency_matrix(g,type = "both"))
  axesprobas = authorsProbas(V(g)$name,8,axes,c(year));teamprobas = authorsProbas(V(g)$name,6,teams,c(year))
 interaxes = t(axesprobas)%*%adjacency%*%axesprobas
 for(axe in axes){if(sum(interaxes[axe,])>0){rate=sum(interaxes[axe,axe])/sum(interaxes[axe,]);
 ccoefs=append(ccoefs,rate);types=append(types,"axes");years=append(years,year);ccoefnames=append(ccoefnames,paste0(axe,year))}}
 interteams = t(teamprobas)%*%adjacency%*%teamprobas
 for(team in teams){if(sum(interteams[team,])>0){rate=sum(interteams[team,team])/sum(interteams[team,]);
 ccoefs=append(ccoefs,rate);types=append(types,"equipes");years=append(years,year);ccoefnames=append(ccoefnames,paste0(team,year))}}
}
names(ccoefs)<-ccoefnames
```

```{r include = TRUE}
g=ggplot(data.frame(ccoef=ccoefs,year=years,type=types),aes(x=year,y=ccoef,colour=type,group=type))
g+geom_point()+geom_smooth()
```


## Distributions of author interdisciplinarities (team and axes)

```{r include = TRUE}
interdisc=c();years=c();types=c()
for(year in allyears){
  show(year)
  load(paste0('res/graph_',year,'.RData'));n=length(V(g)$name)
  axesprobas = authorsProbas(V(g)$name,8,axes,c(year));teamprobas = authorsProbas(V(g)$name,6,teams,c(year))
  interdisc=append(interdisc,1-rowSums(axesprobas^2))
  interdisc=append(interdisc,1-rowSums(teamprobas^2))
  years=append(years,rep(year,2*n));types=append(types,c(rep("axes",n),rep("equipes",n)))
}
```

```{r include = TRUE}
g=ggplot(data.frame(interdisc=interdisc,year=as.character(years),type=types),aes(x=interdisc,colour=year))
g+geom_density()+facet_wrap(~type)
```



Second order interdisc (collaborators)

```{r include = TRUE}
interdisc=c();years=c();types=c()
for(year in allyears){
  show(year)
  # load graph
  load(paste0('res/graph_',year,'.RData'))
  adjacency = as.matrix(as_adjacency_matrix(g,type = "both"))
  axesprobas = authorsProbas(V(g)$name,8,axes,c(year))
  teamprobas = authorsProbas(V(g)$name,6,teams,c(year))
  axesinter = adjacency%*%axesprobas;axesinter=t(apply(axesinter,1,function(r){r/sum(r)}))
  teaminter = adjacency%*%teamprobas;teaminter=t(apply(teaminter,1,function(r){r/sum(r)}))
  interdisc=append(interdisc,1-rowSums(axesinter^2))
  interdisc=append(interdisc,1-rowSums(teaminter^2))
  years=append(years,rep(year,nrow(axesinter)+nrow(teaminter)));types=append(types,c(rep("axes",nrow(axesinter)),rep("equipes",nrow(teaminter))))
}
```

```{r include = TRUE}
g=ggplot(data.frame(interdisc=interdisc,year=as.character(years),type=types),aes(x=interdisc,colour=year))
g+geom_density()+facet_wrap(~type,scales = "free")
```



